Chemical Tagging Using Minimum Message Length

### 2017-04-19 articles/chemical-tagging/ms.tex
some words

Some progress has been made towards chemical tagging in simplistic experiments, 
but the full extent of chemical tagging has yet to be realised. Indeed, in this 
\article\ we assert that even \emph{the full extent of the problem} of chemical 
tagging has yet to be realised. 
% Some other comments about this?

\subsection*{A Probabilistic Approach to Chemical Tagging}
\ref{section:the-problem}

What is needed to make things work? We outline the 12 disciples:

@@

@@

\section{The Model}
\ref{sec:the-model}

@@

\begin{equation}
    f(y_n|\bm{\mu},\bm{\Sigma}) 
        = \frac{1}{\sqrt{(2\pi)^d|\bm{\Sigma}|}}
          \exp{\left[-\frac{1}{2}(\bm{y} - \bm{\mu})^\intercal\bm{\Sigma}^{-1}(\bm{y} - \bm{\mu})\right]}
\end{equation}

@@

\section{Introducing MML}
% MML formalism

@@

>>>>

Some progress has been made towards chemical tagging in simplistic experiments, 
but the full extent of chemical tagging has yet to be realised. Indeed, in this 
\article\ we assert that even \emph{the full extent of the problem} of chemical 
tagging has yet to be realised. Before reviewing the existing literature on
what has been done with chemical tagging, we will outline the necessary
components to perform chemical tagging.

\subsection{The Problem: What's needed for chemical tagging?}
\label{section:the-problem}

Given a set of stars that have been observed spectroscopically, and detailed
chemical abundances have been derived, the following issues will arise when
trying to perform chemical tagging:

@@

Previous approaches have tried to solve separate components of chemical tagging (e.g. clustering). Some notes on this.

These approaches are not consistent, because the problems are intertwined (e.g., clustering, latent factors).

@@

\section{The Model}
\label{sec:the-model}

@@

\begin{equation}
    f(\bm{y}|\bm{\mu},\bm{\Sigma}) 
        = \frac{1}{\sqrt{(2\pi)^d|\bm{\Sigma}|}}
          \exp{\left[-\frac{1}{2}(\bm{y} - \bm{\mu})^\intercal\bm{\Sigma}^{-1}(\bm{y} - \bm{\mu})\right]}
\end{equation}

@@

\subsection{Minimum Message Length}
\label{sec:mml}

Background to MML, etc.

We want to minimize the message length.

Link to Bayesian evidence, other things, etc.

In order to minimize the full message length, we must specify the priors on
our parameters.

@@

\subsection{Expected Fisher information}

Computing the expected Fisher information requires the second order partial
derivatives of the negative log-likelihood $-\mathcal{L}(\bm{y}|\bm{\mu},\bm{\Sigma})$.
This can be tricky for mixtures -- explain why.
We approximate the determinant of of the Fisher information matrix,
$|\mathcal{F}(\bm{\mu},\bm{\Sigma})|$ as the product of $|\mathcal{F}\left(\bm{\mu}\right)|$ and $|\mathcal{F}\left(\bm{\Sigma}\right)|$ \citep{Oliver_1996,Roberts_1998}.

% Equation for log-likelihood of the data:
% \begin{equation}
% \mathcal{L}\left(\bm{y}|\bm{\mu},\bm{\Sigma}) = -\frac{Nd}{2}\log{\left(2\pi\right)} - \frac{N}{2}\log{|\bm{\Sigma}|} - \frac{1}{2}\Sigma_{i=1}^{N}(\bm{y}_i - \bm{mu})^\intercal{}\Sigma^{-1}(\bm{y}_i - \bm{\mu})
%\end{equation}

Taking the second derivative of the log-likelihood of the data with respect to $\bm{\mu}$:

\begin{equation}
\frac{\delta\mathcal{L}}{\delta\bm{\mu}} = \Sigma_{i=1}^{N}\bm{C}^{-1}(\bm{y}_i - \bm{\mu})
\end{equation}

\begin{equation}
\frac{\delta^2\mathcal{L}}{\delta\bm{\mu}^2} = N\bm{C}^{-1}
\end{equation}

Computing the Fisher information of the covariance matrix $|\mathcal{F}\left(\bm{C}\right)|$ is harder. For a full covariance matrix with off-diagonal terms, the total
number of free parameters required is $d(d+1)/2$.
\citep{Magnus_1988} derived an analytical expression:
% Dwyer_1967, Bozdogan_1990? Drton_2009?
% TODO: This derivation is a nightmare to find

\begin{equation}
|\mathcal{F}\left(\bm{C}\right) = N^\frac{d(d+1)}{2}2^{-d}|\bm{C}|^{-(d+1)}
\end{equation}

Giving the approximate determinant of the Fisher information $|\mathcal{F}\left(\bm{\mu},\bm{C}\right)|$ as:

\begin{eqnarray}
  |\mathcal{F}\left(\bm{\mu},\bm{C}\right)| & \approx & |\mathcal{F}\left(\bm{\mu}\right)|\cdot|\mathcal{F}\left(\bm{C}\right)| \\
  & \approx & N^{d}|\bm{C}|^{-1}N^\frac{d(d+1)}{2}2^{-d}|\bm{C}|^{-(d+1)} \\
  |\mathcal{F}\left(\bm{\mu},\bm{C}\right)| & \approx & N^\frac{d(d+3)}{2}2^{-d}|\bm{C}|^{-(d+2)}
\end{eqnarray}

A common method to estimate the parameters of a Gaussian distribution is by
maximum likelihood, where the log-likelihood $\mathcal{L}$ is maximized. In
order to do this, one computes the gradient of the log-likelihood function
with respect to the parameters and solves the resulting equations.

- Encoding the message.

- Process of expectation and maximization for each step.
